Notes for week of 10/4/19 - 10/11/19

### 2021-12-23 notes/weekly_notes/10.11.19.NotesForWeek.tex

Combining Corollary~\ref{cor:mixing_time_PPR} and Theorem~\ref{thm:acl_3} leads to an upper bound on $\Phi(p)$.
\begin{theorem}
	\label{thm:conductance_ppr}
	Let $C$ be a set satisfying
	\begin{itemize}
		\item $\vol(C) \leq \frac{2}{3}\vol(G)$,
		\item $\abs{C} \geq \frac{d_{\max}}{d_{\min}}$, and
		\item $\frac{20\Phi(C)}{1 + 10\Phi(C)} + \frac{d_{\max}}{2d_{\min}^2} \leq \frac{1}{10}$.
	\end{itemize}
Set $\alpha = 10\Phi(C)$. Then, for any $v \in C_{\alpha}$,
\begin{equation*}
\Phi(p(\alpha,\chi_v)) \leq \sqrt{1600\left\{\log\left(\frac{m}{d_{\min}^2}\right) + \log 20\right\} \Phi(C)}
\end{equation*}
\end{theorem}

@@

To apply Theorem~\ref{thm:mixing_time_PPR_contrapositive} to the specific case of $G$ a neighborhood graph, we will need estimates on the degree and volume graph functionals. These estimates will rely on certain regularity properties of $\mathcal{X}$ and $f$. In particular, we will assume
\begin{itemize}
	\label{asmp:regularity_conditions}
	\item There exist $\lambda_{\min}$ and $\lambda_{\max}$ such that for any $x \in \mathcal{X}$:
	\begin{equation*}
	0 < \lambda_{\min} < f(x) < \lambda_{\max} < \infty.
	\end{equation*}
	\item There exists some $a > 0$ such that for any $r >0 $ and any $x \in \mathcal{X}$,
	\begin{equation*}
	\nu(B(x,r) \cap \mathcal{X}) \geq \frac{r^d}{a},
	\end{equation*}
	where $\nu(\cdot)$ is Lebesgue measure on $\Reals^d$ and $B(x,r)$ is the ball centered at $x$ of radius $r$. 
\end{itemize}
We collect our bounds on graph functionals in the following Lemma.
\begin{lemma}
	\label{lem:graph_functional_concentration}
	Suppose $\mathcal{X}$ and $f$ satisfy the regularity properties. Fix $\mathcal{S} \subseteq \mathcal{X}$, and write $\mathcal{S}[X] = \mathcal{X} \cap \set{x_1,\ldots,x_n}$.  For any $\delta \in (0,1)$, suppose
	\begin{equation}
	\label{eqn:graph_functional_concentration_1}
	n \geq a \frac{(1 + \delta)}{(1 - \delta)^2} \cdot \frac{\lambda_{\max}}{\lambda_{\min}} \max\set{\frac{1}{\Pbb(\mathcal{S})}, \frac{10a}{(1-\delta)\lambda_{\min}\nu_dr^d}}.
	\end{equation}
	Then there exists constants $c_1,c_2$ independent of $n$ such that following statements hold with probability at least $1 - c_1n\exp\{-c_2n\}$:
	\begin{itemize}
		\item $\abs{\mathcal{S}[X]} \geq \frac{d_{\max}}{d_{\min}}$, 
		\item $\frac{d_{\max}}{2d_{\min}^2} < \frac{1}{20}$. 
	\end{itemize}
	If additionally
	\begin{equation}
	\label{eqn:graph_functional_concentration_2}
	\mathbb{E}[\vol(\mathcal{S}[X])] \leq 2\frac{(1 - \delta)}{(1 + \delta)} \mathbb{E}[\vol((\mathcal{X}\setminus\mathcal{S})[X])]
	\end{equation}
	then $\vol(\mathcal{S}[X]) \leq \frac{2}{3} \vol(G)$ with probability at least $1 - c_3n\exp\{-c_4n\}$, where $c_3,c_4$ are constants independent of $n$. 
\end{lemma}

Together Lemma~\ref{lem:graph_functional_concentration} and Theorem~\ref{thm:mixing_time_PPR_contrapositive} imply the following result.
\begin{corollary}
	\label{cor:conductance_ppr_neighborhood_graph}
	Fix $\mathcal{S} \subset \mathcal{X}$, and $\delta \in (0,1)$. Suppose 
	\begin{itemize}
		\item $\mathcal{X}$ and $f$ satisfy the regularity properties,
		\item  $\mathcal{S}$ satisfies \eqref{eqn:graph_functional_concentration_2} for any $r \in (0,\infty)$, and
		\item The sample size $n$ satisfies \eqref{eqn:graph_functional_concentration_1},
		\item The normalized cut $\Phi(\mathcal{S}[X]) \leq \frac{1}{200}$.
	\end{itemize}.
	Set $\alpha = \Phi(\mathcal{S}[X])$. For any $v \in \mathcal{S}[X]_{\alpha}$,
	\begin{equation*}
	\Phi(p(\alpha,\chi_v)) \leq \sqrt{1600\left\{2\log\left(\frac{a}{(1 - \delta)\lambda_{\min}\nu_d r^d}\right) + \log 20\right\} \Phi(\mathcal{S}[X])}
	\end{equation*}
	with probability at least $1 - cn\exp\set{-cn}$. 
\end{corollary}
The important takeaway is that $\Phi(p) \lesssim \sqrt{\Phi(\mathcal{S}[X])}$ (where we write $a_n \lesssim b_n$ when there exists constant $c$ such that $a_n \leq c b_n$ for all $n$.) In particular, we have eliminated the factor of $\log(\vol(G))$ which made upper bounds derived from Theorem~\ref{thm:acl_2} facile as $n,\vol(G) \to \infty$.

@@

To show a lower bound for density clustering using PPR, we exhibit a hard case: that is, a distribution $\Pbb$ for which PPR is unlikely to recover a density cluster. Let $\mathcal{C}_0$, $\mathcal{C}_1$, and $\mathcal{C}_2$ be rectangles in $\Reals^2$, 
\begin{equation*}
\mathcal{C}_0 = \left[-\frac{3\sigma}{2}, -\frac{\sigma}{2}\right] \times \left[-\frac{\rho}{2}, \frac{\rho}{2}\right], \quad \mathcal{C}_1 = \mathcal{C}_0 + (\sigma,0), \quad \mathcal{C}_2 = \mathcal{C}_0 + (2\sigma,0)
\end{equation*}
and let $\mathbb{P}$ be the mixture distribution
\begin{equation*}
\mathbb{P} = \frac{1 - \epsilon}{2} \Psi_1 + \frac{1 - \epsilon}{2} \Psi_2 + \frac{\epsilon}{2} \Psi_0
\end{equation*}
where $\Psi_m$ is the uniform distribution over $\mathcal{C}_m$ for $m = 0,1,2$. 
The density function $f$ of $\Pbb$ is clearly
\begin{equation*}
f(x) = \frac{1}{\rho\sigma}\left(\frac{1 - \epsilon}{2}\1(x \in \mathcal{C}_1) + \frac{1 - \epsilon}{2}\1(x \in \mathcal{C}_2) + \frac{\epsilon}{2}\1(x \in \mathcal{C}_0)  \right)
\end{equation*}
so that for any $\epsilon < \lambda < (1 - \epsilon)/2$, $\mathbb{C}_{f}(\lambda) = \set{\mathcal{C}_1, \mathcal{C}_2}$.

@@

\begin{theorem}
	Let \textcolor{red}{$\alpha = 10 \Phi_{n,r}(\mathcal{L}[X])$} and $r = \sigma/4$. Suppose $\frac{r}{\sigma} < \frac{1}{6400}$. Then, there exists a set $C_{\alpha}$ with $\abs{C_{\alpha} \cap \mathcal{C}_1[X]} \geq \abs{\mathcal{C}_1[X]}/6$ such that for any $v \in C_{\alpha}$, the minimum conductance sweep cut $\widehat{C}$ of $p(\alpha,\chi_v)$,
	\begin{equation*}
	\frac{\abs{\widehat{C} \triangle \mathcal{C}_1[X]}}{n} \geq 1 - c \frac{1}{\epsilon^2} \sqrt{\log\left(\frac{\rho \sigma}{\epsilon\sigma^d}\right)\frac{\sigma}{\rho}}
	\end{equation*}
	with probability at least $1 - c_1 n \exp\set{-c_2n}$, where $c_1,c_2$ are constants which do not depend on $n$.
\end{theorem}
\begin{proof}
	Let $\mathcal{L}$ be the bottom half of the rectangle $\mathcal{X}$,
	\begin{equation*}
	\mathcal{L} = \left[-\frac{3\sigma}{2}, \frac{3\sigma}{2}\right] \times \left[-\frac{\rho}{2},\frac{\rho}{2}\right].
	\end{equation*}
	We will take for granted that with probability at least $1 - c_1 n \exp\set{-c_2n}$, the following two statements are true
	\begin{enumerate}
		\item $\Phi(\mathcal{L}[X]) \leq 32 \frac{r}{\rho}$, and
		\item For any set $A \subset X$,
		\begin{equation}
		\label{eqn:conductance_C1}
		\Phi(A) \geq \frac{\epsilon^2 r}{\sigma}\left(1 - \frac{\abs{\widehat{C} \triangle \mathcal{C}_1[X]}}{n(1 - \epsilon)}\right)
		\end{equation}
	\end{enumerate}
	It can be shown that $\mathcal{L}$ and $\mathcal{X},f$ satisfy all of the conditions of Corollary~\ref{cor:conductance_ppr_neighborhood_graph}. Therefore, choosing $\alpha = 10 \Phi_{n,r}(\mathcal{L}[X])$, there exists a set $C_{\alpha}$ with $\vol(C_{\alpha}) \geq \vol(\mathcal{L}[X])/2$ such that 
	\begin{equation*}
	\Phi(\widehat{C}) \leq c \sqrt{\log\left(\frac{\rho \sigma}{\epsilon\sigma^d}\right) \Phi(\mathcal{L}[X])} \leq c \sqrt{\log\left(\frac{\rho \sigma}{\epsilon\sigma^d}\right) \frac{r}{\rho}}.
	\end{equation*}
	Combining this inequality with \eqref{eqn:conductance_C1}, we have
	\begin{equation*}
	\frac{\epsilon^2 r}{\sigma}\left(1 - \frac{\abs{\widehat{C} \triangle \mathcal{C}_1[X]}}{n(1 - \epsilon)}\right) \leq c \sqrt{\log\left(\frac{\rho \sigma}{\epsilon\sigma^d}\right) \frac{r}{\rho}}.
	\end{equation*}
	Plugging in $r = \sigma/4$, and solving for $\frac{\abs{\widehat{C} \triangle \mathcal{C}_1[X]}}{n}$, we obtain the desired result.
\end{proof}

@@

\subsection{Proof of Theorem~\ref{thm:conductance_ppr}.}
Since $\alpha = 10\Phi(C)$ and $v \in C_{\alpha}$, by Theorem~\ref{thm:acl_3},
\begin{equation*}
p(\alpha,\chi_v)(C) \geq \frac{9}{10}.
\end{equation*}
This inequality along with the assumption $\vol(C) \leq \frac{2}{3}\vol(G)$ implies that $p(\alpha,\chi_v)(C) - \frac{\vol(C)}{\vol(G)} \geq \frac{1}{5}$. Since we assume $\abs{C} \geq \frac{d_{\max}}{d_{\min}}$, the hypothesis of Theorem~\ref{thm:mixing_time_PPR_contrapositive} is satisfied with $\delta = 1/5$, and we have
\begin{equation*}
\Phi(p(\alpha,\chi_v)) \leq \sqrt{\frac{160\Phi(C)\left\{\log\left(\frac{m}{d_{\min}^2}\right) + \log\left(\frac{2}{\delta'}\right)\right\}}{\delta'}}
\end{equation*}
Finally, we assume $\frac{20\Phi(C)}{1 + 10\Phi(C)} + \frac{d_{\max}}{2d_{\min}^2} \leq \frac{1}{10}$ which implies that
\begin{equation*}
\delta' = \delta - \frac{20\alpha}{1 + 10\alpha} + \frac{d_{\max}}{2d_{\min}^2} \geq \frac{1}{10} 
\end{equation*}
completing the proof of the theorem.
\section{Technical results.}

@@

\subsection{Concentration Inequalities.}

@@

>>>>

Combining Corollary~\ref{cor:mixing_time_PPR} and Theorem~\ref{thm:acl_3} leads to an upper bound on $\Phi(p)$.
\begin{theorem}
	\label{thm:conductance_ppr}
	Let $C$ be a set satisfying
	\begin{itemize}
		\item $\vol(C) \leq \frac{2}{3}\vol(G)$,
		\item $\abs{C} \geq \frac{d_{\max}}{d_{\min}}$, and
		\item $\frac{20\Phi(C)}{1 + 10\Phi(C)} + \frac{d_{\max}}{2d_{\min}^2} \leq \frac{1}{10}$.
	\end{itemize}
Suppose $10\Phi(C) \leq \alpha \leq 20\Phi(C)$. Then, there exists a subset $C_{\alpha} \subset C$ with $\vol(C_{\alpha}) \geq \vol(C)/2$ such that for any $v \in C_{\alpha}$,
\begin{equation*}
\Phi(p(\alpha,\chi_v)) \leq \sqrt{3200\left\{\log\left(\frac{m}{d_{\min}^2}\right) + \log 20\right\} \Phi(C)}
\end{equation*}
\end{theorem}

@@

To apply Theorem~\ref{thm:conductance_ppr} when $G$ is a random geometric graph, we will need estimates on the degree and volume graph functionals. These estimates will rely on certain regularity properties of $\mathcal{X}$ and $f$. In particular, we will assume
\begin{enumerate}[(A)]
	\item 
	\label{asmp:regularity_condition_1}
	There exist $\lambda_{\min}$ and $\lambda_{\max}$ such that for any $x \in \mathcal{X}$:
	\begin{equation*}
	0 < \lambda_{\min} < f(x) < \lambda_{\max} < \infty.
	\end{equation*}
	\item
	\label{asmp:regularity_condition_2}
	There exists some $a > 0$ such that for any $r >0 $ and any $x \in \mathcal{X}$,
	\begin{equation*}
	\nu(B(x,r) \cap \mathcal{X}) \geq \frac{\nu_d r^d}{a},
	\end{equation*}
	where $\nu(\cdot)$ is Lebesgue measure on $\Reals^d$ and $B(x,r)$ is the ball centered at $x$ of radius $r$. 
\end{enumerate}
We collect our bounds on graph functionals in the following Lemma. Let $\mathcal{S} \subset \mathcal{X}$.
\begin{lemma}
	\label{lem:graph_functional_concentration}
	Let $X = \set{x_1,\ldots,x_n}$ be sampled independently from $\Pbb$, and let $G := G_{n,r}$ be a neighborhood graph over $X$. Suppose $\mathcal{X}$ and $f$ satisfy the regularity properties \ref{asmp:regularity_condition_1} and \ref{asmp:regularity_condition_2}. Then, for any $\delta \in (0,1)$ and any $r \in (0,\infty)$, there exists numbers $c_1 := c_1(f,\mathcal{X},a)$ and $c_2 := c_2(f,\mathcal{X},a)$ independent of the sample size $n$ such that each of the following bounds hold with probability at least $1 - c_1n\exp\set{-c_2n}$:
	\begin{itemize}
		\item $(1 - \delta) \lambda_{\min} \frac{\nu_dr^d}{a} n \leq d_{\min} \leq d_{\max} \leq (1 + \delta) \lambda_{\min} \frac{\nu_dr^d}{a}n$, 
		\item $(1 - \delta) \mathbb{P}(\mathcal{S}) n \leq \abs{\mathcal{S}[X]} \leq (1 - \delta) \mathbb{P}(\mathcal{S}) n$,
		\item $(1 - \delta) \mathbb{E}[\vol(\mathcal{S}[X])] \leq \vol(\mathcal{S}[X]) \leq (1 + \delta) \mathbb{E}[\vol(\mathcal{S}[X])]$,
		\item $(1 - \delta) \mathbb{E}[\mathrm{cut}(\mathcal{S}[X])] \leq \mathrm{cut}(\mathcal{S}[X]) \leq (1 + \delta) \mathbb{E}[\mathrm{cut}(\mathcal{S}[X])]$,
	\end{itemize}
\end{lemma}
The bounds on $d_{\min}$ and $d_{\max}$ follow from standard reasoning, in which we i) use the regularity conditions \ref{asmp:regularity_condition_1} and \ref{asmp:regularity_condition_2} to upper and lower bound the expected degree $\mathbb{E}(d(x_i))$ over all $x_i$, ii) apply Bernstein's inequality to obtain bounds on the deviation $\abs{d(x_i) - \mathbb{E}d(x_i)}$ for a given $i$, and  iii) apply a union bound. Bernstein's inequality also leads to a bound on the deviation of $\abs{S[X]}$ from its mean. Finally, the functionals $\vol(\mathcal{S}[X])$ and $\mathrm{cut}(\mathcal{S}[X])$ are U-statistics and therefore by Hoeffding's inequality concentrate around their respective expectations.

Lemma~\ref{lem:graph_functional_concentration} implies that when the sample size is sufficiently large, the conditions of Theorem~\ref{thm:conductance_ppr} will be met. We therefore obtain the following upper bound on the normalized cut of PPR computed on a neighborhood graph.
\begin{corollary}
	\label{cor:conductance_ppr_neighborhood_graph}
	Fix $\mathcal{S} \subset \mathcal{X}$, radius $r \in (0,\infty)$, and $\delta \in (0,1/7]$. Suppose $\mathcal{X}$ and $f$ satisfy the regularity properties \ref{asmp:regularity_condition_1} and \ref{asmp:regularity_condition_2}, that
	\begin{equation}
	\label{eqn:graph_functional_concentration_1}
	n \geq a \frac{(1 + \delta)}{(1 - \delta)^2} \cdot \frac{\lambda_{\max}}{\lambda_{\min}} \max\set{\frac{1}{\Pbb(\mathcal{S})}, \frac{10a}{(1-\delta)\lambda_{\min}\nu_dr^d}},
	\end{equation}
	and
	\begin{equation}
	\label{eqn:graph_functional_concentration_2}
	\mathbb{E}[\vol(\mathcal{S}[X])] \leq 2\frac{(1 - \delta)}{(1 + \delta)} \mathbb{E}[\vol((\mathcal{X}\setminus\mathcal{S})[X])],
	\end{equation}
	and that $\Phi_{\Pbb}(\mathcal{S}) \leq \frac{1}{200(1 + \delta)}$. Set $\alpha = 15\Phi_{\Pbb}(\mathcal{S})$. Then, there exists a set $\mathcal{S}[X]_{\alpha} \subset \mathcal{S}[X]$ with $\vol(\mathcal{S}[X]_{\alpha}) \leq \vol(\mathcal{S}[X])/2$, such that for any $v \in \mathcal{S}[X]_{\alpha}$,
	\begin{equation*}
	\Phi(p(\alpha,\chi_v)) \leq \sqrt{3200\left\{2\log\left(\frac{a}{(1 - \delta)\lambda_{\min}\nu_d r^d}\right) + \log 20\right\} \Phi(\mathcal{S}[X])}
	\end{equation*}
	with probability at least $1 - c_1n\exp\set{-c_2n}$.
\end{corollary}
The important takeaway is that $\Phi(p) \lesssim \sqrt{\Phi(\mathcal{S}[X])}$ (where we write $a_n \lesssim b_n$ when there exists constant $c$ such that $a_n \leq c b_n$ for all $n$.) In particular, we have eliminated the factor of $\log(\vol(G))$ which made upper bounds derived from Theorem~\ref{thm:acl_2} facile as the sample size $n$ and the volume $\vol(G) \to \infty$.

@@

To show a lower bound for density clustering using PPR, we exhibit a hard case: that is, a distribution $\Pbb$ for which PPR is unlikely to recover a density cluster. Let $\mathcal{C}_0$, $\mathcal{C}_1$, and $\mathcal{C}_2$ be rectangles in $\Reals^2$, 
\begin{equation*}
\mathcal{C}_0 = \left[-\frac{3\sigma}{2}, -\frac{\sigma}{2}\right] \times \left[-\frac{\rho}{2}, \frac{\rho}{2}\right], \quad \mathcal{C}_1 = \mathcal{C}_0 + (\sigma,0), \quad \mathcal{C}_2 = \mathcal{C}_0 + (2\sigma,0)
\end{equation*}
and let $\mathbb{P}$ be the mixture distribution
\begin{equation*}
\mathbb{P} = \frac{1 - \epsilon}{2} \Psi_1 + \frac{1 - \epsilon}{2} \Psi_2 + \frac{\epsilon}{2} \Psi_0
\end{equation*}
where $\Psi_m$ is the uniform distribution over $\mathcal{C}_m$ for $m = 0,1,2$. 
The density function $f$ of $\Pbb$ is simply
\begin{equation*}
f(x) = \frac{1}{\rho\sigma}\left(\frac{1 - \epsilon}{2}\1(x \in \mathcal{C}_1) + \frac{1 - \epsilon}{2}\1(x \in \mathcal{C}_2) + \frac{\epsilon}{2}\1(x \in \mathcal{C}_0)  \right)
\end{equation*}
so that for any $\epsilon < \lambda < (1 - \epsilon)/2$, $\mathbb{C}_{f}(\lambda) = \set{\mathcal{C}_1, \mathcal{C}_2}$.

@@

\begin{theorem}
	Let $\alpha = 15\Phi_{\Pbb}(\mathcal{L})$ and $r = \frac{1}{4}\sigma$. Suppose $\sigma < \frac{\rho}{1600}$. Then, there exists a set $C_{\alpha}$ with $\abs{C_{\alpha} \cap \mathcal{C}_1[X]} \geq \abs{\mathcal{C}_1[X]}/6$ such that for any $v \in C_{\alpha}$, the minimum conductance sweep cut $\widehat{C}$ of $p(\alpha,\chi_v)$ satisfies
	\begin{equation*}
	\frac{\Delta(\widehat{C},\mathcal{C}_1[X])}{n} \geq 1 - c \frac{1}{\epsilon^2} \sqrt{\log\left(\frac{\rho \sigma}{\epsilon\sigma^d}\right)\frac{\sigma}{\rho}}
	\end{equation*}
	with probability at least $1 - c_1 n \exp\set{-c_2n}$, where $c, c_1,c_2$ are constants which do not depend on $n$.
\end{theorem}
\begin{proof}
	Let $\mathcal{L}$ be the bottom half of the rectangle $\mathcal{X}$,
	\begin{equation*}
	\mathcal{L} = \left[-\frac{3\sigma}{2}, \frac{3\sigma}{2}\right] \times \left[-\frac{\rho}{2},\frac{\rho}{2}\right].
	\end{equation*}
	We will take for granted that with probability at least $1 - c_1 n \exp\set{-c_2n}$, the following two statements are true
	\begin{enumerate}
		\item $\Phi_{n,r}(\mathcal{L}[X]) \leq 32 \frac{r}{\rho}$, and
		\item For any set $A \subset X$,
		\begin{equation}
		\label{eqn:conductance_C1}
		\Phi(A) \geq \frac{\epsilon^2 r}{\sigma}\left(1 - \frac{\abs{\widehat{C} \triangle \mathcal{C}_1[X]}}{n(1 - \epsilon)}\right)
		\end{equation}
	\end{enumerate}
	It can be shown that $\mathcal{L}$ and $\mathcal{X},f$ satisfy all of the conditions of Corollary~\ref{cor:conductance_ppr_neighborhood_graph}. Therefore, choosing $\alpha = 10 \Phi_{n,r}(\mathcal{L}[X])$, there exists a set $C_{\alpha}$ with $\vol(C_{\alpha}) \geq \vol(\mathcal{L}[X])/2$ such that 
	\begin{equation*}
	\Phi(\widehat{C}) \leq c \sqrt{\log\left(\frac{\rho \sigma}{\epsilon\sigma^d}\right) \Phi(\mathcal{L}[X])} \leq c \sqrt{\log\left(\frac{\rho \sigma}{\epsilon\sigma^d}\right) \frac{r}{\rho}}.
	\end{equation*}
	Combining this inequality with \eqref{eqn:conductance_C1}, we have
	\begin{equation*}
	\frac{\epsilon^2 r}{\sigma}\left(1 - \frac{\abs{\widehat{C} \triangle \mathcal{C}_1[X]}}{n(1 - \epsilon)}\right) \leq c \sqrt{\log\left(\frac{\rho \sigma}{\epsilon\sigma^d}\right) \frac{r}{\rho}}.
	\end{equation*}
	Plugging in $r = \sigma/4$, and solving for $\frac{\abs{\widehat{C} \triangle \mathcal{C}_1[X]}}{n}$, we obtain the desired result.
\end{proof}

@@

\subsection{Proof of Theorem~\ref{thm:conductance_ppr}.}
Since $\alpha \geq 10\Phi(C)$ and $v \in C_{\alpha}$, by Theorem~\ref{thm:acl_3},
\begin{equation*}
p(\alpha,\chi_v)(C) \geq \frac{9}{10}.
\end{equation*}
This inequality along with the assumption $\vol(C) \leq \frac{2}{3}\vol(G)$ implies that $p(\alpha,\chi_v)(C) - \frac{\vol(C)}{\vol(G)} \geq \frac{1}{5}$. Since we assume $\abs{C} \geq \frac{d_{\max}}{d_{\min}}$, the hypothesis of Theorem~\ref{thm:mixing_time_PPR_contrapositive} is satisfied with $\delta = 1/5$, and we have
\begin{equation*}
\Phi(p(\alpha,\chi_v)) \leq \sqrt{\frac{320\Phi(C)\left\{\log\left(\frac{m}{d_{\min}^2}\right) + \log\left(\frac{2}{\delta'}\right)\right\}}{\delta'}}
\end{equation*}
Finally, we assume $\frac{20\Phi(C)}{1 + 10\Phi(C)} + \frac{d_{\max}}{2d_{\min}^2} \leq \frac{1}{10}$ which implies that
\begin{equation*}
\delta' = \delta - \frac{20\alpha}{1 + 10\alpha} + \frac{d_{\max}}{2d_{\min}^2} \geq \frac{1}{10} 
\end{equation*}
completing the proof of the theorem.

\subsection{Proof of Corollary~\ref{cor:conductance_ppr_neighborhood_graph}.}

Since we assume $\mathcal{X}$ and $f$ satisfy the regularity conditions \ref{asmp:regularity_condition_1} and \ref{asmp:regularity_condition_2}, the inequalities in Lemma~\ref{lem:graph_functional_concentration} hold with probability at least $1 - n c_5 \exp\set{-c_6n}$. We will condition on these inequalities throughout.

We now verify that the conditions required to apply Theorem~\ref{thm:conductance_ppr} are met. The first two results in Lemma \ref{lem:graph_functional_concentration} imply that $\frac{d_{\max}}{d_{\min}} \lesssim \abs{S[X]}$ and $\frac{d_{\max}}{d_{\min}^2} \lesssim 1$; the lower bound on $n$ given in \eqref{eqn:graph_functional_concentration_1} ensures that these bounds hold exactly.

Since we additionally assume that $\Phi_{\Pbb}(\mathcal{S}) \leq \frac{1 - \delta}{200(1 + \delta)}$, the third and fourth results in Lemma \ref{lem:graph_functional_concentration} imply the third condition in Theorem~\ref{thm:conductance_ppr} is also met. Finally, setting $\alpha = 15\Phi_{\Pbb}(\mathcal{S})$, these results imply that
\begin{equation*}
10 \Phi(S)\leq 15\frac{(1 - \delta)}{1 + \delta} \Phi(S)\leq \alpha \leq 15\frac{(1 + \delta)}{1 - \delta} \Phi(S) \leq 20 \Phi(S)
\end{equation*}
where the first and last inequalities follow since $\delta \leq 1/7$. We may therefore apply Theorem~\ref{thm:conductance_ppr} and obtain that
\begin{equation*}
\Phi(p(\alpha,\chi_v)) \leq \sqrt{3200\left\{\log\left(\frac{\vol(G)}{d_{\min}^2}\right) + \log 20\right\} \Phi(C)}
\end{equation*}
Using the (potentially loose) bound $\vol(G) \leq n^2$, Corollary~\ref{cor:conductance_ppr_neighborhood_graph} follows from the first result in Lemma~\ref{lem:graph_functional_concentration}.

\section{Technical results.}

@@

\subsection{Concentration Inequalities.}
\label{subsec:concentration}

@@

\section{Old Stuff.}
\subsection{Proof of Lemma~\ref{lem:graph_functional_concentration}.}
Each statement in Lemma~\ref{lem:graph_functional_concentration} is a standard consequence of the concentration results of \ref{subsec:concentration}. For each $i = 1,\ldots,n$, conditional on $x_i$ the degree functional $d(x_i)$ follows a binomial distribution $d(x_i) \sim \textrm{Bin}(n - 1, \Pbb(B(x_i,r)))$. Moreover by the regularity properties, for any $x \in \mathcal{X}$,
\begin{equation*}
\lambda_{\min} \frac{\nu_d r^d}{a} \leq \Pbb(B(x,r)) \leq \lambda_{\max} \nu_d r^d.
\end{equation*}
Therefore, for any $\delta > 0$, applying Bernstein's inequality and a union bound yields
\begin{equation}
\label{eqn:graph_functional_concentration_pf1}
\lambda_{\min} (1 - \delta) \frac{\nu_d r^d}{a} n \leq d_{\min} \leq d_{\max} \leq  \lambda_{\max} (1 - \delta) \nu_d r^d n
\end{equation}
with probability at least $1 - 2 n\exp\set{-\frac{\delta^2 n \lambda_{\min} \nu_d r^d}{a\left(1 + \frac{\delta}{3}\right)}}$. Additionally $\abs{S[X]} \sim \mathrm{Bin}(n, \Pbb(S))$, and therefore
\begin{equation}
\label{eqn:graph_functional_concentration_pf2}
(1 - \delta) \Pbb(\mathcal{S}) n \leq \abs{S[X]}
\end{equation}
with probability at least $1 - \exp\set{-\frac{\delta^2n\Pbb(\mathcal{S})}{1 + \frac{\delta}{3}}}$.

The inequalities \eqref{eqn:graph_functional_concentration_pf1} and \eqref{eqn:graph_functional_concentration_pf2} imply that with high probability $\frac{d_{\max}}{d_{\min}} \lesssim \abs{S[X]}$ and $\frac{d_{\max}}{d_{\min}^2} \lesssim 1$; the lower bound on $n$ given in \eqref{eqn:graph_functional_concentration_1} ensures that these bounds hold exactly.

Finally, the statement $\vol\bigl(\mathcal{S}[X]\bigr) \leq \frac{2}{3}\vol(G)$ is equivalent to $\vol\bigl(\mathcal{X}\setminus\mathcal{S}[X]\bigr) \geq \frac{1}{2}\vol\bigl(\mathcal{S}[X]\bigr)$. For any set $\mathcal{A} \subseteq \Reals^d$, the volume $\vol\bigl(\mathcal{A}[X]\bigr)$ is a $U$-statistic, since it can be written as
\begin{equation*}
\vol\bigl(\mathcal{A}[X]\bigr) = \sum_{i = 1}^{m} \sum_{j \neq i} \underbrace{\frac{\1(x_i \in \mathcal{S}) \vee \1(x_j \in \mathcal{S})}{2} \1(\norm{x_i - x_j} \leq r)}_{h_\mathcal{A}(x_i,x_j)}
\end{equation*}
Let $p_{h_\mathcal{A}} = \mathbb{E}[h_\mathcal{A}(X_1,X_2)]$ for $X_1$ and $X_2$ independent random variables with distribution $\Pbb$. By Hoeffding's inequality for $U$-statistics we have
\begin{equation}
\label{eqn:graph_functional_concentration_pf3}
\vol\bigl(\mathcal{S}[X]\bigr) \leq (1 + \delta)\mathbb{E}\left[\vol(\mathcal{S}[X])\right], \quad \vol\bigl(\mathcal{S}[X]\bigr) \leq (1 - \delta)\mathbb{E}\left[\vol(\mathcal{X}\setminus\mathcal{S}[X])\right]
\end{equation}
with probability at least $1 - 2\exp\set{-\frac{\delta^2 (n - 1)^2 \min\{p_{h_\mathcal{S}}, p_{h_{\mathcal{X}\setminus{\mathcal{S}}}}\}}{4}}$. If \eqref{eqn:graph_functional_concentration_pf3} is satisfied, then by~\eqref{eqn:graph_functional_concentration_2} the inequality $\vol\bigl(\mathcal{X}\setminus\mathcal{S}[X]\bigr) \geq \frac{1}{2}\vol\bigl(\mathcal{S}[X]\bigr)$ holds, and therefore $\vol\bigl(\mathcal{S}[X]\bigr) \leq \frac{2}{3}\vol(G)$.
